---
title: "Stats 306, Fall 2022, Homework 3"
author: "Jasmine Liu, shanyu"
date: "Due October 22, 11:59PM EST"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Question 1

Let's look at some water quality data from the Backbay National Wildlife Refuge:

> Back Bay National Wildlife Refuge is located in the southeastern corner of the City of Virginia Beach. The refuge was established in 1938 to protect and provide habitat for migrating and wintering waterfowl. Diverse habitats, including beachfront, freshwater marsh, dunes, shrub-scrub and upland forest are home to hundreds of species of birds, reptiles, amphibians, mammals and fish.

```{r}
bbwater <- read_csv("../data/BKB_WaterQualityData_2020084.csv")
dim(bbwater)
```
### Q 1.1

Provide a short summary of the data. How many columns are there? What kinds of data do we have (what classes of R data)? How many columns are quantitative? How many are nominal? Are there any dates or times?

**There are 17 columns there. We have both quantitative data and nominal data. 10 columns are quantitative. Yes, there are three columns --"Read_Date, Time(24:00), and Year".**
summarize(bbwater)
```{r}
bbwater
dim(bbwater)[2]

```

### Q 1.2

Use `group_by` or `count` to find the number of observations at each `Site_Id`. Update the `bbwater` to remove any sites that have fewer than 5 observations.
```{r}

groupby_Site <- group_by(bbwater,Site_Id)|>
  summarize(num_observation=n())
groupby_Site

at_least_5<- filter(groupby_Site,num_observation>5)
at_least_5

filter(bbwater,Site_Id %in% at_least_5$Site_Id)
  
```


### Q 1.3

Create an ECDF plot for the `Year` variable.

What is this plot telling you about measurements made before the late 1980s? Remember the refuge was started in 1938. Do any of these measurements seem out of place? Remove them from the `bbwater` table. Make a second ECDF after removing these probably erroneous entries.
** Measurements made before the late 1980s could be considered as none, except very a few occurred around 1900. There is a measurements done around 1900 seems out of place.**
```{r}
ggplot(bbwater,aes(x=Year))+ stat_ecdf()+geom_histogram(bins=40)
exclude_error<-filter(bbwater,Year>1950)
exclude_error
ggplot(exclude_error,aes(x=Year))+stat_ecdf()+geom_histogram(bins=35)
```

### Q 1.4

Here's a little code that puts all of the numeric data into a table with two columns: `name` (which column) and `value` (the observed value for that variable).

```{r}
bbnumeric <- select(bbwater, where(is.numeric)) |> pivot_longer(everything())
bbnumeric
```

Use this data to make a facet plot that shows a histogram plots for all the variables. Use the `scales = "free"` option to `facet_wrap`. Try a few different number of bins to pick a good number.
```{r}
ggplot(bbnumeric,aes(x=value))+geom_histogram(bins=50)+facet_wrap(~name,scales="free")
```

After seeing the small multiples plot, select a variable to generate histogram where you carefully select the number of bins.   
```{r echo=FALSE}
air_temp<- filter(bbnumeric,name==("AirTemp (C)"))
air_temp
ggplot(air_temp,aes(x=value))+geom_histogram(bins=60)
```

## Question 2

### Q 2.1

Create a histogram for `Water Temp (?C)`. Try several numbers of bins to get a useful plot. Would you say that there is one clear typical value for this plot or would you say it contains more than one cluster of values?

**I think there is a clear typical value for this plot.**
```{r}
Water_temp<-filter(bbnumeric,name==('Water Temp (?C)'))
Water_temp
ggplot(Water_temp,aes(x=value))+geom_histogram(bins=70,binwidth = 1)
```

### Q 2.2

Create a facet plot for `Water Temp (?C)` that breaks out a histogram by `Site_Id`.

Do you think the multiple clusters could be explained by different locations having different temperatures?
**I don't think the temperature ranges for these sites are different enough. Acutally these sites have quite similar temperature range from 0 to  about 30 degree Celsius.**
```{r}
bbwater
Water_temp2<- select(bbwater,Site_Id,'Water Temp (?C)')
Water_temp2
ggplot(Water_temp2,aes(x=`Water Temp (?C)`))+geom_histogram(bins=30,binwidth = 1)+facet_wrap(~Site_Id)
```

### Q 2.3

Look at the `Read_Date` column. Which `lubridate` function could parse this format to a date object? Use it to update the `bbwater` table with `real_date` column. Demonstrate with an ECDF plot of the dates.

**the read_date is in the form "month-day-year", therefore I use mdy."
```{r}
head(bbwater$Read_Date)
#month/day/year %m%d%Y
bbwater2<-mutate(bbwater, real_date=mdy(Read_Date))
bbwater2
ggplot(bbwater2,aes(x=real_date))+stat_ecdf()+geom_histogram(bins=35)
```

### Q 2.4

Now use the `month` function to extra the month from each date. Create a facet plot of histograms of the water temp broken out by month. Does this help explain the multiple clusters we saw in the water temp?
**I think this explains the multiple clusters in the water temp. During summer time like June to Sept, the water temperature has a larger maximum temperature and wider range, and vice versa in winter time.**

```{r}

ggplot(bbwater2,aes(x=`Water Temp (?C)`))+geom_histogram(bins=30,binwidth = 1)+facet_wrap(~month(bbwater2$real_date))
```

## Question 3

### Q 3.1

Create plots of the distributions of `Site_Id` and `Water Depth (m)` (separately). What kinds of data (qualitative, ordinal, nominal) are these two measurements?

**Site_Id is norminal data ad Water depth is quantitative data.**
```{r}
ggplot(bbwater,aes(x=`Site_Id`))+geom_bar()
ggplot(bbwater,aes(x=`Water Depth (m)`))+geom_histogram(bins=40)
```


### Q 3.2

Do you think either of these measurements would benefit from a data a logarithmic transformation? Implement this transformation and update the plot(s).

Did you get any warnings? What happened?

### Q 3.3

Create a plot to show the two variables from 3.1 on one plot. Clearly state if this plot shows the **joint distribution** or a set of **conditional distributions**. Describe the plot and if you see any sites with particularly unusual water levels.


## Question 4

### Q 4.1

Suppose instead of using `tidyverse`'s `read_csv` function we had instead used R's built in `read.csv` function to load our data.

Use this function to load the same water quality data set. What are three differences in how the type of data loaded by `read.csv` can be used compared to the data from `read_csv`.

### Q 4.2

Convert the result of 4.1 to a tibble. Demonstrate by printing out the result.

### Q 4.3

Create two tibbles to recreate the first 4 rows of the table of lectures and topics on our [Syllabus page](https://umich.instructure.com/courses/554168/assignments/syllabus). In the first, use `tibble`. For the second, use `tribble`.

### Q 4.4

Open up the file 'homework3.txt'. Use a `read_` function to load this data. You will probably need to read the documentation to deal with a few problems in the file.


